---
title: "Assignment 1 FML"
author: "Hitaishi Bairapureddy"
date: "2024-02-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Summary
This document contains the analysis of global unemployment data of all countries.

## R Markdown
```{r}
Global_unemployment <- read.csv("C:\\Users\\bhita\\Downloads\\archive\\global_unemployment_data.csv")
View(Global_unemployment)
```
This file has been imported from https://www.kaggle.com/datasets/sazidthe1/global-unemployment-data

```{r}
mode(Global_unemployment$X2024)
```
```{r}
median(Global_unemployment$X2020)
```
```{r}
mean(Global_unemployment$X2020)
```
```{r}
max(Global_unemployment$X2020)
```
```{r}
sd(Global_unemployment$X2020)
```
The above values represent descriptive statistics for selection of quantitative variable and the shows mode, median, mean, maximum value and standard deviation for the quantitative variables.


```{r}
str(Global_unemployment$age_categories)
```
The above values represent categorical descriptive analysis of variables

```{r}
Global_unemployment_transformed <- (Global_unemployment$X2020 - mean(Global_unemployment$X2020)/median(Global_unemployment$X2020))
head(Global_unemployment_transformed)
```
The above represents the transformation of variables.

```{r}
hist(Global_unemployment$X2020)
```
The above represents the graphical representation of histogram
```{r}
x <- Global_unemployment$X2020
y <- Global_unemployment$X2024
plot(x,y, main = "Scatterplot", xlab = "x2020", ylab = "x2024")
```
The above represents the graphical representation of scatterplot and the variables selected are unemployement in 2020 and 2024.


